The Analog Method for Precipitation Prediction: Finding Better Analog Situations at a Sub-Daily Time Step

### 2016-11-24 Moving temporal window - HESS/template.tex
Correction of tenses.

The MTW resulted in a better analogy in terms of the atmospheric circulation, with improved values of the analogy criterion on the whole distribution of analog dates. The improvement was found to grow with the analog ranks due to an accumulation of more similar situations in the selection. Moreover, the improvement is even more important for days with heavy precipitation events, which are generally related to more dynamic atmospheric situations, where timing is more specific, and which are less present in the archive.

@@

However, the MTW requires sub-daily precipitation time series, which are generally available on a shorter period than daily precipitation archives. Thus, two simple temporal disaggregation methods were assessed in order to make longer archives usable with the MTW. The assessed approaches were not successful, which emphasized the need to use time series with an appropriate chronology. These should be available in a near future, either by means of growing observed archives, or by the establishment of precipitation reanalyses through regional modeling. Then, the use of a MTW in the AM should be considered for any application, and in any context, let it be operational forecasting or climate-related studies

@@

Analog methods (AMs) are based on the hypothesis that two relatively similar synoptic situations may produce similar local effects \citep{Lorenz1956, Lorenz1969}. They are part of the statistical downscaling methods \cite{Maraun2010}. They consist in finding past situations that are similar, in terms of the atmospheric circulation or other synoptic predictors, to the target day of interest. Then, the local weather variables of interest (predictand) that were observed at the analog dates are used to construct a probabilistic prediction for the target day \citep{Duband1970, Zorita1999}. Multiple variations of the methods exist \citep[a non exhaustive listing can be found in][]{BenDaoud2016}. AMs are used for operational precipitation forecasting, either in the context of flood forecasting or hydropower production \citep[e.g.,][]{Guilbaud1997, Bontron2005, Hamill2006, Desaint2008a, GarciaHernandez2009b, Bliefernicht2010, Marty2010, Marty2012, Horton2012, Obled2014, Hamill2015, BenDaoud2016}, as well as precipitation downscaling in a climate perspective \citep[e.g.,][]{Radanovics2013, Chardon2014, Dayon2015}. Other applications focus on temperatures \citep{Radinovic1975, Woodcock1980, Kruizinga1983, DelleMonache2013, Caillouet2016}, wind \citep{Gordon1987, DelleMonache2013, DelleMonache2011, Vanvyve2015, Alessandrini2015, Junk2015, Junk2015c}, solar power \citep{Alessandrini2015a, Bessa2015}, snow avalanches \citep{Obled1980, Bolognesi1993}, insolation \citep{Bois1981}, and the trajectories of tropical cyclones \citep{Keenan1981, Sievers2000, Fraedrich2003}.

The spatial transferability of the method was recently analyzed in depth \citep{Chardon2014}. A great advantage of AMs is that they create realistic events, because they were observed, with consistent spatial distribution, as long as the analog dates chosen for a region are the same. The temporal transferability has also been the topic of recent studies, may it be for past or future climates \citep{Dayon2015, Caillouet2016}.

\citet{Ruosteenoja1988} and \citet{Vandendool1994} have put a major effort into analyzing the influence of the archive length on the quality of the analogy. They have highlighted a three-way relationship between the quality of the analogy, the archive length and the size of the spatial domain (of degrees of freedom): errors increase with a bigger domain, but decrease with a longer archive. They demonstrated that it is not possible to find good analogues on the whole northern hemisphere with a 100 years archive (and even for much higher orders of magnitude). For that reason, smaller spatial windows are always considered to look for analogues, and the archive length is maximized.

@@

The MTW allows finding analog situations at different hours of the day, which can also be seen as an inflation of the archive length. However, despite having $x$ time more candidate situations, the quantity of new information is not expected to be as important as a $x$ time longer archive, due to significant autocorrelation between successive situations. Nevertheless, if the MTW can improve the prediction skill of the AM, it means that it does bring some new information from the archive. Thus, the reduction of the archive length necessary to the MTW, due to the reduced availability of sub-daily precipitation time series, is expected to decrease the AM performance, whereas the archive inflation brought by the MTW is expected to contribute to an increase in performance.

Other possibilities exist in order to increase the prediction skill of AMs. A classical approach is to add new predictors or new successive levels of analogy \citep[see e.g.][]{Horton2012a, BenDaoud2016, Caillouet2016}. AMs can also be combined with other methods \citep[see e.g.][]{Chardon2014}. Another possibility is to a global optimization technique, such as genetic algorithms, in order to better optimize the method and to add new options \citep{Horton2016}. However, the MTW technique is not in competition with other approaches, but can be combined with these. Indeed, as most of them rely on the atmospheric circulation in the first level of analogy, the application of a MTW should lead to similar conclusions.

Section \ref{sec:data_methods} presents the context of the present study as well as data and methods, including the proposed MTW technique. The impact of the reduction of the archive and the improvements brought by the MTW are assessed in Sect. \ref{sec:results}. The benefit of introducing a MTW was assessed first in regards to the analogy criterion improvement (Sect. \ref{sec:influence_criteria}), and then in terms of precipitation prediction skill (Sect. \ref{sec:influence_scores}). Finally, the results are discussed in Sect. \ref{sec:discussion}.

@@

The study area is the upper Rh\^{o}ne catchment in Switzerland. Precipitation time series come from 6 automatic weather stations, namely Ulrichen, Zermatt, Visp, Montana, Sion and Aigle (Fig. \ref{fig:map}), which are subject to various meteorological influences \citep{Horton2012}. The data are available at an hourly time step on 25~years (1982 to 2007), and were also obtained at a daily time step on 47~years (1961 to 2008). Due to the low density of weather stations with high temporal resolution and long archives, no spatially aggregated rainfall was processed. The results will thereafter be presented arbitrarily for the Ulrichen station, but are equivalent for all stations.

Synoptic-scale variables, used as predictors, are extracted from the NCEP/NCAR reanalysis I \citep{Kalnay1996} dataset with a 6-hourly temporal resolution, 17 pressure levels, and a spatial resolution of 2.5\degree. This dataset is now relatively old, but still often used (see discussion in Sect. \ref{sec:old_reanalysis}).

@@

The predictand, here a 24~h precipitation total (classically from 6~h~UTC to 6~h~UTC the next day), results from meteorological situations that are continuous transitions between changing circulation patterns. Thus, the predictor data, extracted from reanalysis datasets, are considered at different hours of the day. Based on \citet{Bontron2005}, geopotential heights are thus compared at 1000~hPa (Z1000) at 12~h~UTC and 500~hPa (Z500) at 24~h~UTC. The time of the day at which the predictors are selected was found by \citet{Bontron2004} to have a significant influence.

@@

Applying the MTW technique involve the need of sub-daily precipitation timeseries. However, such data are not available on long periods. As the AM benefits from long archives, it would be profitable to be able to apply the MTW to a longer archive (here, 1961 to 2008). Thus, in order to reconstruct a longer archive of moving 24~h-totals, two simple disaggregation approaches of the daily time series were assessed.

The first technique is a proportional distribution, where the observed daily precipitation is considered constant during the measuring period (6~h~UTC to 6~h~UTC the next day). It then consists in allocating proportional parts of the original daily time series into a new moving average of 24~h-totals (Fig. \ref{fig:illustration_disaggregation}).

The second approach aims at getting closer to the chronology of the actual precipitation by relying on some informative variable during the reconstruction procedure. The best proxy variable is identified by means of a correlation analyses (on non-zero values) with the 6-hourly precipitation time series on the period 1982-2007. Precipitable water and relative humidity (at 1000~hPa, 925~hPa or 850~hPa) were assessed at the four points surrounding the catchment, and by allowing a time lapse between both series in order to take into account the significant distance separating the weather stations and the reanalysis grid point. Data from the NCEP/NCAR reanalysis 1 (Sect. \ref{sec:data}) was used in this perspective, despite their rough resolution. Then, once the best proxy has been selected, its temporal profile is used in order to disaggregate the precipitation time series on the long archive period.

@@

The calibration is thus done in successive steps, on a limited number of parameters. Previously calibrated parameters are generally not reassessed (except for the number of analogs). More advanced techniques, such as using Genetic Algorithms \citep{Horton2016}, exist, but are out of the scope of the present study.

@@

The performance loss resulting from a reduction of a 47~years archive (1961 to 2008, corresponding to the daily time series) to 25~years (1982 to 2007, corresponding to the hourly time series, see Sect. \ref{sec:mtw}) was assessed with the original method, without MTW. Both 2Z (Table \ref{table:method_2Z}) and 2Z-2MI (Table \ref{table:method_2Z-2MI}) methods were considered. The AM parameters were calibrated on the original archive (see resulting analog numbers in Tables \ref{table:analog_nb}) and will be used thereafter.

The impact of the archive reduction can be seen in Fig. \ref{fig:plots_CRPSS_2Z} for the 2Z method and in Fig. \ref{fig:plots_CRPSS_2Z-2MI} for the 2Z-2MI method. As expected, a loss of performance can be observed for each station, except for that of Aigle, which seems relatively indifferent to this change. This loss is globally significant, with up to -1.89 points for Visp and the 2Z method.

@@

When searching for analogs in the first level of analogy, such as on the geopotential heights in the 2Z method (Sect. \ref{sec:analog_method}), there are 4 times more candidates (even though not fully independent) with the MTW than before (Sect. \ref{sec:mtw}). Figure \ref{fig:changes_S1_analogs} shows the changes in the distributions of the analogy criterion (S1) for the $1^{st}$, $5^{th}$, $20^{th}$ and $40^{th}$ analog rank at the Ulrichen station on the whole calibration period (1961-2008 -- we can use the full period here, as we do not consider precipitation at this stage), due to introduction of the MTW. The shape of the distributions of the conventional approach and the MTW are similar, but the values of the analogy criterion are now reduced (shifted to the left), and therefore better. For analogs with higher ranks (e.g. $20^{th}$ or $40^{th}$), the difference between the two distributions is larger than for the first ranks, which means that the improvement increases with the rank of the analogs.

The improvements of the S1 criterion with the rank of the analogs are summarized in Fig. \ref{fig:changes_S1}, which shows (top) quantiles of the S1 criterion for the conventional method and the MTW, and (bottom) quantiles of the relative reduction (meaning improvement) due to the MTW. This confirms that all quantiles are similarly reduced (S1 distributions keep their shape), and that this improvement is constantly increasing from the first to the last analog (Fig. \ref{fig:changes_S1} bottom).

@@

The impact of the MTW on the analogy criterion has been analyzed per precipitation classes. The results are summarized in Fig. \ref{fig:changes_S1_precip_threshold} by the median reduction of S1 for days with precipitation (organized into classes) between two thresholds. The number of cases per class being reduced, the curves are not as smooth as in previous analyzes. It is nevertheless clear that the improvement tends to increase on days with higher precipitation. This is true for all our stations.

@@

Atmospheric dynamics varies greatly from one season to another, which reflects on the performance of the AM that is generally lower between June and August \citep{Bliefernicht2010}. The effect of the MTW on the S1 criterion per season is presented in Fig. \ref{fig:changes_S1_seasons}. Differences between the seasons are substantial. The improvements are greater for winter than summer.

@@

When adding the second level of analogy of the 2Z-2MI method (Table \ref{table:method_2Z-2MI}), the number of candidate situations did not increase, as they are subsampled in the previously $N_{1}$ selected analogs, but their dates have changed. In contrast to the AM on the atmospheric circulation only, both a reduction or an increase of the RMSE analogy criterion values are possible here with the MTW compared to the static approach. Indeed, Fig. \ref{fig:changes_RMSE} shows an almost insignificant improvement of the RMSE values. Unlike for the first level of analogy, the relative improvement of the RMSE values are distributed relatively symmetrically around zero, with improvements and losses of the same amplitude. Once again, the results for the other stations are similar.

@@

The new performance scores (CRPSS) are provided in Fig. \ref{fig:plots_CRPSS_2Z} and \ref{fig:plots_CRPSS_2Z-2MI} for the 2Z and 2Z-2MI methods respectively, on the reduced archive (1982--2007). The gains, relatively to the static approach on the same archive, range from 0.57 to 2.14 points for the 2Z method and from 1.53 to 2.20 points for the 2Z-2MI method.

@@

The S1 criterion was previously found to be improved to a greater extent for higher precipitation values (Sect. \ref{sec:influence_precip}). The changes in terms of performance scores are now also assessed per precipitation classes. Fig. \ref{fig:changes_CRPS_precip_threshold} synthesizes these differences for the Ulrichen station, other stations having the same behavior. The performance score is improved for days with higher precipitation after the introduction of the MTW. On the contrary, non-rainy days and small precipitation values are not improved.

@@

After recalibrating, changes in the AM parameters can be observed for both 2Z and 2Z-2MI methods. Among these, the zonal extent of the spatial windows of the circulation analogy tends to decrease (not shown). More significantly, the optimal number of analogs increases after introducing the MTW, of a considerable order of magnitude: 25~\% to 83~\% for the 2Z method and 20~\% 67~\% for the 2Z-2MI method (Table \ref{table:analog_nb}). The number of analogs of the first analogy level of the 2Z-2MI method even reached three times its previous value for the Visp station.

The values of the CRPSS scores for both methods (Fig. \ref{fig:plots_CRPSS_2Z} and \ref{fig:plots_CRPSS_2Z-2MI}) have significantly increased after recalibration. When analyzing the change in performance per precipitation classes, the results (not shown) are very similar to what was observed in Sect. \ref{sec:improvement_CRPSS_precip_threshold}, with a slight performance increase of small precipitation values that can be observed at the expense of higher amounts.

@@

On the first level of analogy, the median of the S1 values reduction (Fig. \ref{fig:changes_S1} bottom) starts approximately at 5~\% for the first analog and reaches more than 10~\% for the last one. This increase of improvements with the analog rank can be explained by the accumulation of better analogs in the distribution. The curve representing the minimum improvement is quickly slightly superior to zero, meaning that the criteria have been improved on most analogs for every day of the calibration period. All other stations have a similar improvement of the S1 criterion, both in terms of distribution shape and amplitude. The introduction of the MTW allows finding better analog situations in the first level of analogy, resulting in a selection of days with more similar atmospheric circulation.

When adding a second level of analogy on moisture variables, the criteria values (RMSE) are not improved by the MTW as the number of candidates is not higher (still the $N_{1}$ days selected in the first level of analogy). However, this result of a globally null improvement of the RMSE values does not mean that the 2Z-2MI method cannot be improved by the MTW. It means that after the selection of the analog situations in terms of the synoptic circulation, the new candidate dates do not allow to find better analogs in terms of moisture. However, the selected dates have changed in the first level of analogy, and thus also in the final selection, which can potentially improve the prediction skill.

@@

Finding better analog situations does not necessarily imply better skills to predict the precipitation. This had to be assessed and the MTW was found to improve the skill of the precipitation prediction for both the 2Z and 2Z-2MI methods (Fig. \ref{fig:plots_CRPSS_2Z} and \ref{fig:plots_CRPSS_2Z-2MI}). The improvement of the performance is not very large, but is nevertheless substantial. Moreover, it requires no additional predictor. With the introduction of the MTW, the performance loss related to the reduction of the archive length is compensated. It is here of the same order of magnitude as if the length of the archive doubled. Note that despite the number of candidate situations are four times higher, the gains seem to be lower than for a quadrupled archive length. The likely reason is that an actual longer archive contains more atmospheric situations that might have been less observed on a shorter period. Moreover, high correlations between sub-daily circulation patterns are expected. The MTW thus does not create four times more independent data, but the increase in information is nevertheless substantial. In this regards, a reanalysis dataset with a higher temporal resolution might not improve the performance significantly than a 6-hourly. On the other hand, reducing the MTW to a 12-hourly time step might reduce redundancy in the archive, but might also reduce the chance of finding better analog situations, as circulation patterns can evolve substantially in 12~hours.

@@

Both the analogy criteria (Sect. \ref{sec:influence_precip}) and the performance scores (Sect. \ref{sec:improvement_CRPSS_precip_threshold}) are improved to a greater extent for heavier precipitation events. This is likely due to the fact that higher precipitation events are a consequence of atmospheric conditions with a greater dynamics, such as weather disturbances, which have a well marked temporal evolution. Indeed, the position of the driving elements such as the low-pressure center and the fronts change significantly during a day. Moreover, these situations are less numerous than anticyclonic situations, which makes it less likely to find very good analogs at the same time of the day. We can therefore expect to improve more significantly these situations with a greater dynamics when introducing a MTW, as better matches to the target situation may be found. On the opposite, days with a low dynamics of the atmospheric circulation, such as anticyclonic situations, will not be radically improved by the introduction of the MTW.

@@

An analysis of the selected hours for the geopotential predictor seems to confirm this assumption (Fig. \ref{fig:hours_selection_per_season}). It was found that the new choice of the temporal window in winter, when using the MTW approach, is well balanced between the 4 options. This means a change of 75\% of the analogs selection compared to the conventional approach, which improves the circulation analogy. On the contrary, the summer months have a preference for the initial temporal window (Z500 24~h \& Z1000 12~h), likely due to more pronounced diurnal effects which reduces the potential for improvement of the criteria. Other seasons are between these two extremes, which is consistent with their respective improvements. This seasonal effect was observed for each station in a very similar way, and generally even with a slightly larger amplitude than for Ulrichen.

@@

After the recalibration of the AMs with the MTW, the optimal analog numbers were significantly higher than for the original methods (see Sect. \ref{sec:recalibration} and Fig. \ref{table:analog_nb}). As we saw in Fig. \ref{fig:changes_S1}, the improvement of the S1 criterion grows along with the rank of the analog, which shows an accumulation of better analog situations in the distributions. It seems thus profitable to widen the selection of analogs in order to keep also some whose rank has increased, as they appear to be also relevant to predict precipitation values. The number of good analogs is thus globally increased.

A higher number of analogs generally means, with an archive of fixed length, a poorer analogy. Indeed, when the choice of the predictors or the parameters are improved, leading to a better prediction, the optimal number of analog situations decreases. However, when the length of the archive increases, the optimal number of analogs increases too, for a better performance, up to a certain threshold (\citealt{Bontron2004}, see also Fig. 7 of \citealt{Hamill2006a}). The observed increases in the number of analog situations with the MTW result in better performance skills for a given method, as it can be seen as an inflation of the archive. However, if new relevant predictors were added to the method, the number of analogs would then decrease.

>>>>

The MTW resulted in a better analogy in terms of the atmospheric circulation, with improved values of the analogy criterion on the whole distribution of analog dates. The improvement was found to grow with the analog ranks due to an accumulation of more similar situations in the selection. Moreover, the improvement was even more important for days with heavy precipitation events, which are generally related to more dynamic atmospheric situations, where timing is more specific, and which are less present in the archive.

@@

However, the MTW requires sub-daily precipitation time series, which are generally available on a shorter period than daily precipitation archives. Thus, two simple temporal disaggregation methods were assessed in order to make longer archives usable with the MTW. The assessed approaches were not successful, which emphasized the need to use time series with an appropriate chronology. These should be available in a near future, either by means of growing observed archives, or by the establishment of precipitation reanalyses through regional modeling. Then, the use of a MTW in the AM can be considered for any application, and in any context, let it be operational forecasting or climate-related studies

@@

Analog methods (AMs) are based on the hypothesis that two relatively similar synoptic situations may produce similar local effects \citep{Lorenz1956, Lorenz1969}. They are part of statistical downscaling methods \cite{Maraun2010}. They consist in finding past situations that are similar, in terms of the atmospheric circulation or other synoptic predictors, to the target day of interest. Then, the local weather variables of interest (predictand) that were observed at the analog dates are used to construct a probabilistic prediction for the target day \citep{Duband1970, Zorita1999}. Multiple variations of the methods exist \citep[a non exhaustive listing can be found in][]{BenDaoud2016}. AMs are used for operational precipitation forecasting, either in the context of flood forecasting or hydropower production \citep[e.g.,][]{Guilbaud1997, Bontron2005, Hamill2006, Desaint2008a, GarciaHernandez2009b, Bliefernicht2010, Marty2010, Marty2012, Horton2012, Obled2014, Hamill2015, BenDaoud2016}, as well as precipitation downscaling in a climate perspective \citep[e.g.,][]{Radanovics2013, Chardon2014, Dayon2015}. Other applications focus on temperatures \citep{Radinovic1975, Woodcock1980, Kruizinga1983, DelleMonache2013, Caillouet2016}, wind \citep{Gordon1987, DelleMonache2013, DelleMonache2011, Vanvyve2015, Alessandrini2015, Junk2015, Junk2015c}, solar power \citep{Alessandrini2015a, Bessa2015}, snow avalanches \citep{Obled1980, Bolognesi1993}, insolation \citep{Bois1981}, and the trajectories of tropical cyclones \citep{Keenan1981, Sievers2000, Fraedrich2003}.

The spatial transferability of the method is analyzed in \citet{Chardon2014}. A great advantage of AMs is that they create realistic events, as they are based on observed situations, with consistent spatial distribution, as long as the analog dates chosen for a region are the same. The temporal transferability has also been the topic of recent studies, may it be for past or future climates \citep{Dayon2015, Caillouet2016}.

\citet{Ruosteenoja1988} and \citet{Vandendool1994} analyze the influence of the archive length on the quality of the analogy. They highlight a three-way relationship between the quality of the analogy, the archive length and the size of the spatial domain (of degrees of freedom): errors increase with a bigger domain, but decrease with a longer archive. They demonstrate that it is not possible to find good analogues on the whole northern hemisphere with a 100 years archive (and even for much higher orders of magnitude). For that reason, smaller spatial windows are always considered to search for analogues, and the archive length is maximized.

@@

The MTW allows finding analog situations at different hours of the day, which can also be seen as an inflation of the archive length. However, despite having $x$ time more candidate situations, the quantity of new information is not expected to be as important as a $x$ time longer archive, due to significant correlation between successive situations. Nevertheless, if the MTW can improve the prediction skill of the AM, it means that it does bring some new information from the archive. Thus, if the reduction of the archive length necessary to the MTW, due to the reduced availability of sub-daily precipitation time series, is expected to decrease the AM performance, the archive inflation brought by the MTW is expected to contribute to an increase in performance.

Other possibilities exist in order to increase the prediction skill of AMs. A classical approach is to add new predictors or new successive levels of analogy \citep[see e.g.][]{Horton2012a, BenDaoud2016, Caillouet2016}. AMs can also be combined with other methods \citep[see e.g.][]{Chardon2014}. Another possibility is to use a global optimization technique, such as genetic algorithms, in order to better optimize the method and to add new options \citep{Horton2016}. However, the MTW technique is not in competition with other approaches, but can be combined with these. Indeed, as most of them rely on the atmospheric circulation in the first level of analogy, the application of a MTW should lead to similar conclusions.

Section \ref{sec:data_methods} presents the context of the study as well as data and methods, including the proposed MTW technique. The impact of the reduction of the archive and the improvements brought by the MTW are assessed in Sect. \ref{sec:results}. The benefit of introducing a MTW was assessed first in regards to the analogy criterion improvement (Sect. \ref{sec:influence_criteria}), and then in terms of precipitation prediction skill (Sect. \ref{sec:influence_scores}). Finally, the results are discussed in Sect. \ref{sec:discussion}.

@@

The study area is the upper Rh\^{o}ne catchment in Switzerland. Precipitation time series come from 6 automatic weather stations, namely Ulrichen, Zermatt, Visp, Montana, Sion and Aigle (Fig. \ref{fig:map}), which are subject to various meteorological influences \citep{Horton2012}. The data were available at an hourly time step on 25~years (1982 to 2007), and were also obtained at a daily time step on 47~years (1961 to 2008). Due to the low density of weather stations with high temporal resolution and long archives, no spatially aggregated rainfall was processed. The results will thereafter be presented arbitrarily for the Ulrichen station, but were equivalent for all stations.

Synoptic-scale variables, used as predictors, were extracted from the NCEP/NCAR reanalysis I \citep{Kalnay1996} dataset with a 6-hourly temporal resolution, 17 pressure levels, and a spatial resolution of 2.5\degree. This dataset is now relatively old, but still often used (see discussion in Sect. \ref{sec:old_reanalysis}).

@@

The predictand, here a 24~h precipitation total (classically from 6~h~UTC to 6~h~UTC the next day), results from meteorological situations that are continuous transitions between changing circulation patterns. Thus, the predictor data, extracted from reanalysis datasets, are considered at different hours of the day. Based on \citet{Bontron2005}, geopotential heights are thus compared at 1000~hPa (Z1000) at 12~h~UTC and 500~hPa (Z500) at 24~h~UTC. The time of the day at which the predictors are selected is found by \citet{Bontron2004} to have a significant influence.

@@

It would be profitable to be able to apply the MTW to a longer archive (here, 1961 to 2008), rather than being limited to the reduced period corresponding to the sub-daily precipitation data. Thus, in order to reconstruct a longer archive of moving 24~h-totals, two simple disaggregation approaches of the daily precipitation time series were assessed.

The first technique was a proportional distribution, where the observed daily precipitations were considered constant during the measuring period (6~h~UTC to 6~h~UTC the next day). It then consisted in allocating proportional parts of the original daily time series into a new moving average of 24~h-totals (Fig. \ref{fig:illustration_disaggregation}).

The second approach aimed at getting closer to the chronology of the actual precipitation by relying on some informative variable during the reconstruction procedure. The best proxy variable was identified by means of a correlation analyses (on non-zero values) with the 6-hourly precipitation time series on the period 1982-2007. Precipitable water and relative humidity (at 1000~hPa, 925~hPa or 850~hPa) were assessed at the four points surrounding the catchment, and by allowing a time lapse between both series in order to take into account the significant distance separating the weather stations and the reanalysis grid point. Data from the NCEP/NCAR reanalysis 1 (Sect. \ref{sec:data}) was used in this perspective, despite their rough resolution. Then, once the best proxy has been selected, its temporal profile was used in order to disaggregate the precipitation time series on the long archive period.

@@

The calibration is thus done in successive steps, on a limited number of parameters. Previously calibrated parameters are generally not reassessed (except for the number of analogs). More advanced techniques, such as using genetic algorithms \citep{Horton2016}, exist, but are out of the scope of the present study.

@@

The performance loss resulting from a reduction of a 47~years archive (1961 to 2008, corresponding to the daily time series) to 25~years (1982 to 2007, corresponding to the hourly time series, see Sect. \ref{sec:mtw}) was assessed with the original method, without MTW. Both 2Z (Table \ref{table:method_2Z}) and 2Z-2MI (Table \ref{table:method_2Z-2MI}) methods were considered. The AM parameters were calibrated on the original archive (see resulting analog numbers in Tables \ref{table:analog_nb}) and were be used thereafter.

The impact of the archive reduction can be seen in Fig. \ref{fig:plots_CRPSS_2Z} for the 2Z method and in Fig. \ref{fig:plots_CRPSS_2Z-2MI} for the 2Z-2MI method. As expected, a loss of performance could be observed for each station, except for that of Aigle, which seemed relatively indifferent to this change. This loss was globally significant, with up to -1.89 points for Visp and the 2Z method.

@@

When searching for analogs in the first level of analogy, such as on the geopotential heights in the 2Z method (Sect. \ref{sec:analog_method}), there are 4 times more candidates (even though not fully independent) with the MTW than before (Sect. \ref{sec:mtw}). Figure \ref{fig:changes_S1_analogs} shows the changes in the distributions of the analogy criterion (S1) for the $1^{st}$, $5^{th}$, $20^{th}$ and $40^{th}$ analog rank at the Ulrichen station on the whole calibration period (1961-2008 -- the full period could be used here, as precipitation were not considered at this stage), due to introduction of the MTW. The shape of the distributions of the conventional approach and the MTW were found to be similar, but the values of the analogy criterion were reduced (shifted to the left), and therefore better. For analogs with higher ranks (e.g. $20^{th}$ or $40^{th}$), the difference between the two distributions was larger than for the first ranks, which means that the improvement increased with the rank of the analogs.

The improvements of the S1 criterion with the rank of the analogs are summarized in Fig. \ref{fig:changes_S1}, which shows (top) quantiles of the S1 criterion for the conventional method and the MTW, and (bottom) quantiles of the relative reduction (meaning improvement) due to the MTW. This confirms that all quantiles were similarly reduced (S1 distributions keep their shape), and that this improvement was constantly increasing from the first to the last analog (Fig. \ref{fig:changes_S1} bottom).

@@

The impact of the MTW on the analogy criterion has been analyzed per precipitation classes. The results are summarized in Fig. \ref{fig:changes_S1_precip_threshold} by the median reduction of S1 for days with precipitation (organized into classes) between two thresholds. The number of cases per class being reduced, the curves are not as smooth as in previous analyzes. It is nevertheless clear that the improvement were more important on days with higher precipitation. It was the same for all other stations.

@@

Atmospheric dynamics varies greatly from one season to another, which reflects on the performance of the AM that is generally lower between June and August \citep{Bliefernicht2010}. The effect of the MTW on the S1 criterion per season is presented in Fig. \ref{fig:changes_S1_seasons}. Differences between the seasons were substantial. The improvements were greater for winter than summer.

@@

When adding the second level of analogy of the 2Z-2MI method (Table \ref{table:method_2Z-2MI}), the number of candidate situations did not increase, as they were subsampled in the previously $N_{1}$ selected analogs, but their dates have changed. In contrast to the AM on the atmospheric circulation only, both a reduction or an increase of the RMSE analogy criterion values were possible here with the MTW compared to the static approach. Indeed, Fig. \ref{fig:changes_RMSE} shows an almost insignificant improvement of the RMSE values. Unlike for the first level of analogy, the relative improvement of the RMSE values were distributed relatively symmetrically around zero, with improvements and losses of the same amplitude. Once again, the results for the other stations were similar.

@@

The new performance scores (CRPSS) are provided in Fig. \ref{fig:plots_CRPSS_2Z} and \ref{fig:plots_CRPSS_2Z-2MI} for the 2Z and 2Z-2MI methods respectively, on the reduced archive (1982--2007). The gains, relatively to the static approach on the same archive, ranged from 0.57 to 2.14 points for the 2Z method and from 1.53 to 2.20 points for the 2Z-2MI method.

@@

The S1 criterion was previously found to be improved to a greater extent for higher precipitation values (Sect. \ref{sec:influence_precip}). The changes in terms of performance scores were then also assessed per precipitation classes. Fig. \ref{fig:changes_CRPS_precip_threshold} synthesizes these differences for the Ulrichen station, other stations having the same behavior. The performance score was improved for days with higher precipitation after the introduction of the MTW. On the contrary, non-rainy days and small precipitation values were not improved.

@@

After recalibrating, changes in the AM parameters could be observed for both 2Z and 2Z-2MI methods. Among these, the zonal extent of the spatial windows of the circulation analogy decrease a bit (not shown). More significantly, the optimal number of analogs increased after introducing the MTW, of a considerable order of magnitude: 25~\% to 83~\% for the 2Z method and 20~\% 67~\% for the 2Z-2MI method (Table \ref{table:analog_nb}). The number of analogs of the first analogy level of the 2Z-2MI method even reached three times its previous value for the Visp station.

The values of the CRPSS scores for both methods (Fig. \ref{fig:plots_CRPSS_2Z} and \ref{fig:plots_CRPSS_2Z-2MI}) have significantly increased after recalibration. When analyzing the change in performance per precipitation classes, the results (not shown) were very similar to what was observed in Sect. \ref{sec:improvement_CRPSS_precip_threshold}, with a slight performance increase of small precipitation values that can be observed at the expense of higher amounts.

@@

On the first level of analogy, the median of the S1 values reduction (Fig. \ref{fig:changes_S1} bottom) started approximately at 5~\% for the first analog and reached more than 10~\% for the last one. This increase of improvements with the analog rank can be explained by the accumulation of better analogs in the distribution. The curve representing the minimum improvement is quickly slightly superior to zero, meaning that the criteria have been improved on most analogs for every day of the calibration period. All other stations had a similar improvement of the S1 criterion, both in terms of distribution shape and amplitude. The introduction of the MTW allowed finding better analog situations in the first level of analogy, resulting in a selection of days with more similar atmospheric circulation.

When adding a second level of analogy on moisture variables, the criteria values (RMSE) were not improved by the MTW as the number of candidates was not higher (still the $N_{1}$ days selected in the first level of analogy). However, this result of a globally null improvement of the RMSE values does not mean that the 2Z-2MI method cannot be improved by the MTW. It means that after the selection of the analog situations in terms of the synoptic circulation, the new candidate dates did not allow to find better analogs in terms of moisture. However, the selected dates have changed in the first level of analogy, and thus also in the final selection, which can potentially improve the prediction skill.

@@

Finding better analog situations does not necessarily imply better skills to predict the precipitation. This had to be assessed and the MTW was found to improve the skill of the precipitation prediction for both the 2Z and 2Z-2MI methods (Fig. \ref{fig:plots_CRPSS_2Z} and \ref{fig:plots_CRPSS_2Z-2MI}). Moreover, it required no additional predictor. With the introduction of the MTW, the performance loss related to the reduction of the archive length was compensated. It was here of the same order of magnitude as if the length of the archive doubled. Note that despite the number of candidate situations were four times higher, the gains seem to be lower than for a quadrupled archive length. The likely reason is that an actual longer archive would contain more atmospheric situations that might have been less observed on a shorter period. Moreover, high correlations between sub-daily circulation patterns are expected. The MTW thus did not create four times more independent data, but the increase in information was nevertheless substantial. In this regards, a reanalysis dataset with a higher temporal resolution might not improve the performance significantly more than a 6-hourly dataset. On the other hand, reducing the MTW to a 12-hourly time step might reduce redundancy in the archive, but might also reduce the chance of finding better analog situations, as circulation patterns can evolve substantially in 12~hours.

@@

Both the analogy criteria (Sect. \ref{sec:influence_precip}) and the performance scores (Sect. \ref{sec:improvement_CRPSS_precip_threshold}) were improved to a greater extent for heavier precipitation events. This is likely due to the fact that higher precipitation events are a consequence of atmospheric conditions with a greater dynamics, such as weather disturbances, which have a well marked temporal evolution. Indeed, the position of the driving elements such as the low-pressure center and the fronts change significantly during a day. Moreover, these situations are less numerous than anticyclonic situations, which makes it less likely to find very good analogs at the same time of the day. We can therefore expect to improve more significantly these situations with a greater dynamics when introducing a MTW, as better matches to the target situation may be found. On the opposite, days with a low dynamics of the atmospheric circulation, such as anticyclonic situations, will not be radically improved by the introduction of the MTW.

@@

An analysis of the selected hours for the geopotential predictor seems to confirm this assumption (Fig. \ref{fig:hours_selection_per_season}). It was found that the new choice of the temporal window in winter, when using the MTW approach, is well balanced between the 4 options. This means a change of 75\% of the analogs selection compared to the conventional approach, which improved the circulation analogy. On the contrary, days in summer months had a preference for the initial temporal window (Z500 24~h \& Z1000 12~h), likely due to more pronounced diurnal effects which reduces the potential for improvement of the criteria. Other seasons were between these two extremes, which is consistent with their respective improvements. This seasonal effect was observed for each station in a very similar way, and generally even with a slightly larger amplitude than for Ulrichen.

@@

After the recalibration of the AMs with the MTW, the optimal analog numbers were significantly higher than for the original methods (see Sect. \ref{sec:recalibration} and Fig. \ref{table:analog_nb}). As we saw in Fig. \ref{fig:changes_S1}, the improvement of the S1 criterion grew along with the rank of the analog, which shows an accumulation of better analog situations in the distributions. It seems thus profitable to widen the selection of analogs in order to keep also some whose rank has increased, as they appear to be also relevant to predict precipitation values. The number of good analogs was thus globally increased.

A higher number of analogs generally means, with an archive of fixed length, a poorer analogy. Indeed, when the choice of the predictors or the parameters are improved, leading to a better prediction, the optimal number of analog situations decreases. However, when the length of the archive increases, the optimal number of analogs increases too, for a better performance, up to a certain threshold (\citealt{Bontron2004}, see also Fig. 7 of \citealt{Hamill2006a}). The observed increases in the number of analog situations with the MTW resulted in better performance skills for a given method, as it can be seen as an inflation of the archive. However, if new relevant predictors were added to the method, the number of analogs would then decrease.
